---
title: "Stem_and_Leaf"
output:
  pdf_document: default
  html_notebook: default
---


The stem method in PACKAGE:graphics produces a basic stem and leaf plot. THe method works best on un-skewwed, outlier-free, shorter datasets. 
```{r}
?stem

# "ideal" Example (n = 100) --------------------------------
stem(Nile)
```

The "scale" argument adjusts the bin size. 
```{r}
# bins Example ---------------------------------------------
# scale = 1
stem(faithful$eruptions)

# decreasing the scale decreases the number of bins
stem(faithful$eruptions, scale = 0.5) 

# increasing the scale increases the number of bins
stem(faithful$eruptions, scale = 2)
```

Transformations can be employed to help with skew, but be aware that transformations hinder interpretation.
```{r}
# Skew Example ---------------------------------------------
# Heavy skew
stem(islands)

# Better.
stem(log(islands))
```

The method is not appropriate for very large datasets. Try to summarize the data in some way to reduce the number of entries (e.g., monthly data --> annual data)
```{r}
# Too many entries Example ---------------------------------
# (n = 3177)
stem(sunspot.month)

# (n = 289)
stem(sunspot.year)
```

Additional control is possible with stem.leaf and stem.leaf.backback in the PACKAGE:aplpack
```{r}
require(aplpack)
stem.leaf(Nile)
```

